---
title: "DS project"
author: "Kalide Endale"
date: "12/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(DescTools)
library(modelr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(randomForest)
```

## R Markdown
```{r}
dir01 <- "/Users/Kalide/Documents/Northeastern"
dir02 <-"Introduction to Data Management and Processing/Ames.csv"
path01 <- file.path(dir01,dir02)
data1 <- read_csv(path01)
```

Preliminary results
```{r}
data1 <-subset(data1,select=-c(Order,PID))

group_mean <- aggregate(SalePrice ~ Yr.Sold, data = data1, mean)

graph1<-ggplot(group_mean,aes(x=Yr.Sold,y=SalePrice))+
  geom_point()+geom_smooth();graph1

ols<-lm(SalePrice ~ Year.Built,data1)


summary(ols)

correlation<-cor(data1$SalePrice,data1[, unlist(lapply(data1, is.numeric))])
```


Alley: NA = no alley acess
Pool : NA = no pool
Misc feature: NA = no additional features
Fence: NA = no fence
Fireplace quality: no fireplace
Lot frontage (Linear feet of street connected to property
) = has missing data, but most likely won't use. not an imp. indicator
```{r}
#count number of NA's in file and visualize NAs
missing.values <- data1 %>%
  gather(key = "key", value = "val") %>%
  mutate(is.missing = is.na(val)) %>%
  group_by(key, is.missing) %>%
  dplyr::summarise(num.missing = n()) %>%
  filter(is.missing == T) %>%
  select(-is.missing) %>%
  arrange(desc(num.missing))


#plot the chart of NAs
missing.values %>%
  ggplot() +
  geom_bar(aes(x=reorder(key, +num.missing), y=num.missing, fill = key), 
           stat = 'identity') + 
  coord_flip() +
  labs(x='variable', 
       y="number of missing values", 
       title='Visual of Missing Values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme_minimal()
```


```{r}
attach(data1)
#remove variables we don't need and remove NAs
data2 <- subset(data1,
                select = -c(Pool.QC,Misc.Feature,Alley,Fence,
                            Fireplace.Qu,Lot.Frontage,Garage.Yr.Blt,Garage.Qual,
                            Garage.Finish,Garage.Cond,Bsmt.Exposure,BsmtFin.Type.1,
                            BsmtFin.Type.2))

data2 <- na.omit(data2)
```



```{r}
#turn central air into a dummy variable
data2$Central.Air <- ifelse(data2$Central.Air == "Y", 1,0)
data2$Street <- ifelse(data2$Street == "Pave",1,0)
```

Street - turn into dummy (1 for pavement)
Remove:
Lot shape? Possible remove
Lot config? possible remove
Land contour and lot slope are similar ? possibly remove one
Condition 1 and Condition 2 (proximity to various conditions) Both columns have mostly normal proximity. Possibly remove both
Building Type and House Style are similar. Remove House style
Overall condition:remove, since most of our ind. variables are describing condition of home
Roof style and Roof matl are similar. Remove Roof style because material is better indicator of cost and quality
Mas vnr type and Mas vnr Area can both be removed. We have exterior quality on house as an indicator
Exterior 1st and Exterior 2nd: remove both since we have exterior quality on house as an indicator
BasmtFin SF 1 (Type 1 finished square feet) and BsmtFin Type 2(Rating of basement finished area): remove since square footage of entire home is included 
X1st.Flr.SF and X2nd.Flr.SF: removing both square feet of these variables because we already have living area as an indicator variable
Low quality finished square feet: remove
bedroom.above grd, kitchen above grade will be removed. We will keep total rooms above grade as an indicator
Garage cars and Garage Area: are very similar. We will remove garage area.
3-san porch and Screen porch will be removed. We have open porch sf and enclosed porch sf as indicator variables for proch size
Misc.val (value of miscellaneous feature) - remove. not important

```{r}
data2 <- subset(data2,
                select = -c(Lot.Shape,Lot.Config,Land.Contour,
                            Condition.1,Condition.2,House.Style,
                            Overall.Cond,Roof.Style,Mas.Vnr.Area,
                            Mas.Vnr.Type,Exterior.1st,Exterior.2nd,
                            BsmtFin.SF.1,BsmtFin.SF.2,X1st.Flr.SF,
                            X2nd.Flr.SF,Low.Qual.Fin.SF,Bedroom.AbvGr,
                            Kitchen.AbvGr,Garage.Area,X3Ssn.Porch,
                            Screen.Porch,Misc.Val,Utilities))
data2$Neighborhood <- factor(data2$Neighborhood)
```


#additoinal removal of variables that are not a significant predictor of sales price

#we will remove variables that have no signifcance with SalePrice. 

Anova on variables to see statistacal signficance
#adjusted our p-values using a bonferroni correction with an alpha cutoff of 0.05;
conservative adjustment
strongly safeguard against making any type 1 errors at cost of lower power
reduced our our candidate variables from 45 to 27

```{r}
lm01 <- lm(log10(SalePrice) ~ Overall.Qual, data = data2)

summary(lm01)

#looks like there is one very large residual in the first box plot
data2 %>%
  add_residuals(lm01, "resid") %>%
  ggplot(aes(x=Overall.Qual, group = Overall.Qual)) +
  geom_boxplot(aes(y=resid)) +
  labs(x="Overall Quality", 
       y="Residuals",
       title = "Residuals vs. Overall Quality") +
  theme_minimal()
#qq plot: cleary shows the we have an outlier
data2 %>%
  add_residuals(lm01, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal() +
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantilies", 
       title = "QQ Plot: Standardized residuals" )


#The results are the same when we plot this in a histogram. The histogram looks like a normal distribution and one outlier in the beginning of the graph.
data2 %>%
  add_residuals(lm01, "resid") %>%
  ggplot(aes(x=resid)) +
  geom_histogram(bins=100) +
  labs(x="Residuals",
       title = "Distribution of residuals") +
  theme_minimal()
```

```{r}
#store the residuals of the linear regression as a new column in data2
data2$resid <- resid(lm01)
#Find out what 3 standard deviation is for the distribution and save it
SD3 <- 3*sd(resid(lm01))
#Make a dummy variable column to find out which data points exceed 3 s.d.
data2$exceeds_3sd <- ifelse(abs(data2$resid) > SD3, 1, 0)

#filter the outlier from dataframe2
data2 <- data2 %>%
  filter(data2$exceeds_3sd == 0)
```


```{r}
# remove 1 Neighborhood because it only has one data. Will not allow us to properly partition data
data2 <- data2 %>%
  filter(Neighborhood!='Landmrk')

data2 <- select(data2, -c(resid,exceeds_3sd))

data2
```



```{r}
set.seed(4)
data_part <- resample_partition(data2,
                            p=c(train=0.5,
                                valid=0.25,
                                test=0.25))
data2 <- as.data.frame(data_part$train)
dataValid <- as.data.frame(data_part$valid)
dataTest <- as.data.frame(data_part$test)
data_part
data2
```

```{r}
#remove insignifcant variables
ANOVAsummary <- aov(SalePrice ~., data = data2)

summary(ANOVAsummary)
```

```{r}
library(broom)

data3 <- ANOVAsummary %>%
  tidy() %>%
  mutate(adj.p.value = p.adjust(p.value,
                                method = "bonferroni")) %>%
  filter(adj.p.value < 0.05)
data3
```

#reduced model after ANOVA test using bonferroni
```{r}
data4 <- data2[,which((names(data2) %in% data3$term)== TRUE)]
data4 <- data.frame(data4, data2$SalePrice) %>%
  rename(SalePrice = data2.SalePrice)
data4
```

Correlation for numeric values
```{r}
data4.cor <- data4 %>% 
select_if(is.numeric)
```

correlation<-cor(df03.cor$log10gdp_percapita,df03.cor[, unlist(lapply(df03.cor, is.numeric))])
t(correlation)

#corrletation matrix
```{r fig.width=10, fig.height=10}
data4.cor <- subset(data4.cor)
ggcorrplot::ggcorrplot(cor(data4.cor), tl.cex = 12) 
numeric_cor <- cor(data4.cor$SalePrice,data4.cor[, unlist(lapply(data4.cor,is.numeric))])
t(numeric_cor)

numeric_cor <- as.tibble(t(numeric_cor))
numeric_cor
num_corr <- data.frame(colnames(data4.cor[, unlist(lapply(data4.cor,is.numeric))]),numeric_cor)
num_corr
```

```{r}
data4.cor_catg <- data4 %>% 
select_if(Negate(is.numeric))
```



```{r}
corr_cat <- c()
for(i in colnames(data4.cor_catg)){
corr <- UncertCoef(data4.cor_catg[[i]], data2$SalePrice, direction = "column")
corr_cat=c(corr_cat,corr)

}

correlation_categorical <- data.frame(colnames(data4.cor_catg), corr_cat)
correlation_categorical[order(-corr_cat),]
```


```{r}
colnames(num_corr) <- c('Features','Correlation')
colnames(correlation_categorical) <- c('Features','Correlation')

#final correlation plot
data_final_corr <- rbind(num_corr, correlation_categorical) %>%
  mutate(abs_correlation = abs(Correlation)) %>%
  arrange(desc(abs_correlation))
data_final_corr

```



Variables we will choose from that have the best correlation with SalePrice
Overall quality
Gr.Liv.Area
GarageCars
Total.Bsmt.Sf
Year.Built
Year.Remod.Add
Neighborhood (the only categorical variable that had a decent correlation with saleprice)
Lot Area

```{r}
dataFinal <- select(data4, Overall.Qual, Gr.Liv.Area, Garage.Cars,
                    Total.Bsmt.SF, Year.Built, Year.Remod.Add,
                    Neighborhood, Lot.Area, SalePrice)

dataFinalValid <- select(dataValid, Overall.Qual, Gr.Liv.Area, 
                         Garage.Cars, Total.Bsmt.SF, Year.Built,
                         Year.Remod.Add, Neighborhood, Lot.Area, SalePrice)

dataFinalTest <- select(dataTest, Overall.Qual, Gr.Liv.Area, 
                         Garage.Cars, Total.Bsmt.SF, Year.Built,
                        Year.Remod.Add, Neighborhood, Lot.Area, SalePrice)
dataFinal
dataFinalValid
dataFinalTest
```


```{r}
rf <- randomForest(log10(SalePrice)~., data=dataFinal, proximity=TRUE)
rf1 <- randomForest(log10(SalePrice) ~., data = data2, proximity = TRUE)
```

```{r}
print(rf)
plot(rf)
rmse(rf, dataFinalValid)
mae(rf, dataFinalValid)

dataFinalValid %>%
  add_residuals(rf, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

#RF against 43 varialbes instead of 9
print(rf1)
plot(rf1)
rmse(rf1, dataValid)
mae(rf1, dataValid)

dataValid %>%
  add_residuals(rf1, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()
```

```{r}
ggplot(data = data1, aes(Yr.Sold, fill = Neighborhood)) +
  geom_bar(show.legend = FALSE) + facet_wrap(~Neighborhood) + 
  theme(axis.text.x = element_text(color = "black", 
                                   size = 7, angle=45,
                                   vjust=.8, hjust=0.8))
```

ggplot(data = data1) + 
         geom_bar(aes(Yr.Sold, fill = Yr.Sold)) + coord_flip() + scale_fill_viridis(discrete = TRUE, option = "A") + theme_minimal()

```{r}
data1 %>%
ggplot(aes(x=SalePrice, group = Yr.Sold, fill = Yr.Sold)) +
geom_histogram(bindwith = 10) + 
theme(axis.text.x = element_text(color = "black", 
                                 size = 6, angle=30,
                                 vjust=.8, hjust=0.8)) + 
scale_fill_viridis(discrete = FALSE, option = "H") + theme_minimal()
```


```{r}
ggplot(dataFinal, aes(x=Overall.Qual, y=log10(SalePrice), 
                      group = Overall.Qual, fill = Overall.Qual)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = FALSE) +
  theme_minimal()
```

```{r}
ggplot(dataFinal, aes(x=Gr.Liv.Area, y=log10(SalePrice))) +
  geom_point() +
  geom_smooth(fill="light blue") +
  geom_smooth(method="lm", color="red", fill="pink") +
  theme_minimal()
```

```{r}
ggplot(dataFinal, aes(x=Garage.Cars, y=log10(SalePrice), group = Garage.Cars)) +
  geom_boxplot(aes(fill = Garage.Cars)) +
  scale_fill_viridis(discrete = FALSE, option = "G")+
  theme_minimal()
```

```{r}
ggplot(dataFinal, aes(x=Total.Bsmt.SF, y=log2(SalePrice))) +
  geom_point() +
  geom_smooth(fill="light blue") +
  geom_smooth(method="lm", color="red", fill="pink") +
  theme_minimal()
```

```{r}
ggplot(dataFinal, aes(x=Year.Built, y=log2(SalePrice))) +
  geom_point() +
  geom_smooth(fill="light blue") +
  geom_smooth(method="lm", color="red", fill="pink") +
  theme_minimal()
```

```{r}
ggplot(dataFinal, aes(x=Year.Remod.Add, y=log10(SalePrice))) +
  geom_point() +
  geom_smooth(fill="light blue") +
  geom_smooth(method="lm", color="red", fill="pink") +
  theme_minimal()
```
scale_colour_viridis_d(option = "plasma")
fill = Overall.Qual)
```{r}
ggplot(dataFinal, aes(x=Gr.Liv.Area, y=mean(log10(SalePrice)), 
                      group = Overall.Qual)) +
  geom_boxplot(aes(fill = Overall.Qual)) +
  scale_fill_viridis(discrete = FALSE, option = "H")  +
  theme_minimal() +
  labs(y = "Average log10(saleprice)")
  theme(axis.text.x = element_text(color = "black", 
                                   size = 9, angle=30,
                                   vjust=.8, hjust=0.8))
  
```

```{r}
makeStep <- function(response, predictors, candidates)
{
  rhs <- paste0(paste0(predictors, collapse="+"), "+", candidates)
  formulas <- lapply(paste0(response, "~", rhs), as.formula)
  rmses <- sapply(formulas,
                  function(fm) rmse(lm(fm, data=dataFinal),
                                    data=dataFinalValid))
  names(rmses) <- candidates
  attr(rmses, "best") <- rmses[which.min(rmses)]
  rmses
}
```

#### Step 1 (no variables):

```{r}
model <- NULL

preds <- ""
cands <- c("Overall.Qual", 
           "Gr.Liv.Area", 
           "Garage.Cars", 
           "Total.Bsmt.SF", 
           "Year.Built",
           "Year.Remod.Add",
           "Neighborhood",
           "Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```

```{r}
preds <- c("Overall.Qual")
cands <- c("Gr.Liv.Area", 
           "Garage.Cars", 
           "Total.Bsmt.SF", 
           "Year.Built",
           "Year.Remod.Add",
           "Neighborhood", "Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```

```{r}
preds <- c("Overall.Qual", "Gr.Liv.Area")
cands <- c("Neighborhood", 
           "Garage.Cars", 
           "Total.Bsmt.SF", 
           "Year.Built",
           "Year.Remod.Add","Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```

```{r}
preds <- c("Gr.Liv.Area", "Overall.Qual", "Neighborhood")
cands <- c("Garage.Cars", 
           "Total.Bsmt.SF", 
           "Year.Built",
           "Year.Remod.Add","Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```

```{r}
preds <- c("Gr.Liv.Area", "Overall.Qual", "Neighborhood",
           "Total.Bsmt.SF")
cands <- c("Garage.Cars","Year.Built","Year.Remod.Add","Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```
```{r}
preds <- c("Gr.Liv.Area", "Overall.Qual", "Neighborhood",
           "Total.Bsmt.SF","Year.Remod.Add")
cands <- c("Garage.Cars","Year.Built","Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```

```{r}
preds <- c("Gr.Liv.Area", "Overall.Qual", "Neighborhood",
           "Year.Built", "Garage.Cars","Year.Built")
cands <- c("Garage.Cars","Lot.Area")
s1 <- makeStep("log10(SalePrice)", preds, cands)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
step_model <- tibble(index=seq_along(model),
                     variable=factor(names(model), levels=unique(names(model))),
                     RMSE=model)

ggplot(step_model, aes(y=RMSE)) +
  geom_point(aes(x=variable)) +
  geom_line(aes(x=index)) +
  labs(title="Stepwise model selection") +
  theme_classic() +
  theme(axis.text.x = element_text(color = "black", 
                                   size = 9, angle=30,
                                   vjust=.8, hjust=0.8))
```

```{r}
library(olsrr)
modelStepwise <- lm(log10(SalePrice) ~ Overall.Qual + Gr.Liv.Area + Neighborhood + Total.Bsmt.SF + Year.Remod.Add + Year.Built, data=dataFinalValid)

rmse(modelStepwise, dataFinalValid)

dataFinalValid %>%
  add_residuals(modelStepwise, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal() +
  coord_cartesian(ylim=c(-.3,.2))

dataFinalValid %>%
add_residuals(modelStepwise, "resid") %>%
ggplot(aes(x=resid)) +
geom_histogram(bins=50, color = "darkblue", fill = "lightblue") +
theme_minimal()

#residual test against variables


dataFinalValid %>%
add_residuals(modelStepwise, "resid") %>%
ggplot(aes(x= Overall.Qual)) +
geom_point(aes(y=resid))+
labs(x="",
     y="Residuals",
     title = "Residuals vs. Overall Quality")
theme_minimal()

dataFinalValid %>%
add_residuals(modelStepwise, "resid") %>%
ggplot(aes(x= Gr.Liv.Area)) +
geom_point(aes(y=resid))+
labs(x="",
     y="Residuals",
     title = "Residuals vs. Above Ground Living Area Square Feet")
theme_minimal()

  
dataFinalValid %>%
add_residuals(modelStepwise, "resid") %>%
ggplot(aes(x=Neighborhood)) +
geom_point(aes(y=resid))+
labs(x="",
     y="Residuals",
     title = "Residuals vs. Neighborhood")
theme_minimal() +
theme(axis.text.x = element_text(color = "black", 
                                 size = 5, angle=45,
                                 vjust=.8, hjust=0.8))

dataFinalValid %>%
  add_residuals(modelStepwise, "resid") %>%
  ggplot(aes(x= Total.Bsmt.SF)) +
  geom_point(aes(y=resid))+
  labs(x="Total Basment Square Feet",
       y="Residuals",
       title = "Residuals vs. Total Basment Square Feet")
  theme_minimal()

dataFinalValid %>%
add_residuals(modelStepwise, "resid") %>%
ggplot(aes(x= Year.Remod.Add)) +
geom_point(aes(y=resid))+
labs(x="Year Remodeled",
     y="Residuals",
     title = "Residuals vs. Year Remodeled")
theme_minimal()  
  
  
dataFinalValid %>%
add_residuals(modelStepwise, "resid") %>%
ggplot(aes(x= Year.Built)) +
geom_point(aes(y=resid))+
labs(x="Year Built",
     y="Residuals",
     title = "Residuals vs. Year Built")
theme_minimal()

```


```{r}
summary(modelStepwise)
```


```{r}
dataFinal <- subset(dataFinal, select = -Neighborhood)
dataFinal$SalePrice <- log10(dataFinal$SalePrice)

dataFinal <- dataFinal %>%
  rename(Year.Remod = Year.Remod.Add)
```


```{r fig.width=8, fig.height=7}


library("GGally")
lowerFn <- function(data, mapping, method = "lm",...) {
p <- ggplot(data = data, mapping = mapping) +
geom_point(colour = "blue") +
geom_smooth(method = method, color = "red",...)
p
}

ggpairs(dataFinal, lower = list(continuous = wrap(lowerFn, method = "lm")),
diag = list(continuous = wrap("densityDiag", colour = "blue")),
upper = list(continuous = wrap("cor", size = 4))) + 
theme(axis.text.x = element_text(color = "black",
                                 size = 8, angle=30, 
                                 vjust=.8, hjust=0.8))
```


```{r}
library(vip)
vip(modelStepwise, aesthetics = list(fill = "purple")) + theme_classic()


```

